---
title: "First Analysis Write-Up"
author: "Yiying Wu (yw3996)"
date: ""
output: 
  pdf_document:
    latex_engine: pdflatex
header-includes:
   - \usepackage{mathptmx} # Times New Roman font or similar
   - \usepackage{setspace} # Package for setting spacing
   - \usepackage[margin=1in]{geometry} # Adjust page margins if needed
   - \doublespacing # Command to apply double spacing
fontsize: 12pt
classoption: doublespace
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(gtsummary)
library(tidyverse)
library(dplyr)
library(knitr)
library(kableExtra)
```

**Note: All codes are available here**: https://github.com/wuyiying2018/food-insecurity

# 1. Introduction


# 2. Methodology
## 2.1 Data sources
The dataset used in the report is the NHANES 2017-March 2020 dataset. This survey is conducted, and the data is collected by the National Center for Health Statistics (NCHS) from 2017 to March 2020. 

The NHANES 2019-2020 field operations were suspended in March 2020 due to the COVID-19 pandemic, resulting in incomplete data collection that is not nationally representative. To address this issue and ensure national representativeness, the partially collected 2019-2020 data were merged with the complete 2017-2018 dataset, creating a combined dataset representative of the U.S. civilian non-institutionalized population before the pandemic. The dataset contains demographic, socioeconomic, dietary, and health-related information. 

The target population is adults over 20 years old (includes 20) in U.S.

## 2.2 Weighting the Survey Data
In NHANES dataset weights are adjusted for selection probability, non-response, and post- stratification to match U.S. population figures. To obtain a valid statistical inference, a domain analysis for the adult subpopulation (age $\ge$ 20) are conducted by using the subset function in the survey package in R.


## 2.3 Multiple Imputation

To handle missing data in our NHANES dataset, we utilize Multiple Imputation (MI) to generate multiple datasets by imputing missing values repeatedly. This method retains the advantages of single imputation, such as consistent analyses and data collector’s knowledge, while also accurately reflecting uncertainty and accounting for imputation error. We implemented it using the MICE (Multivariate Imputation by Chained Equations) package in R, which allows for flexible and efficient imputation of missing values. This ensures that the imputed values are plausible and improves statistical efficiency.

## 2.4 Logistic Regression

Given the binary outcome: food security (Yes/No), we chose logistic regression as our statistical model. Suppose there are $n$ covariates $X_i, i=1,\ ...\ n$, the model can be expressed as:

$$
log(\frac{\pi}{1-\pi})=X\beta = \beta_0+\beta_1X_1+...+\beta_nX_n
$$

- $\pi$ is the risk

- $\frac{\pi}{1-\pi}$ is the odds

- $\beta_0$ is the log odds for $X_i's=0$

- $\beta_i$ is the log odds ratio per unit change of $X_i$, holding all other covariates fixed

In the logistic regression analysis for survey data, I have integrated multiple imputation and complex survey design in NHANES. Each imputed dataset was then individually analyzed with logistic regression models, incorporating essential survey design elements like stratification and weights via the survey package. Lastly, these results are aggregated to derive final estimates using Rubin’s multiple imputation combining rule (Rubin, 2018).

## 2.5 Machine Learning Method

plan to use the `surveyCV` package

some useful links:

https://github.com/ColbyStatSvyRsch/surveyCV

https://stats.stackexchange.com/questions/238141/two-worlds-collide-using-ml-for-complex-survey-data

https://cran.r-project.org/web/packages/surveyCV/index.html


# 3. Results
## 3.1 Descriptive Statistics
\newpage
\begin{singlespace}
```{r echo=FALSE, messgae=FALSE, warning=FALSE}
my_dat <- read_csv("result_files/dat.csv")%>%
  filter(age>=20)

dat <- my_dat %>%
  mutate(
    gender = factor(case_when(
      female == "1" ~ "female",  
      female == "0" ~ "male"),
      levels = c("male", "female")
    ),
    food_security= factor(case_when(
      food_security == "1" ~ "food security",  
      food_security == "0" ~ "food insecurity"),
      levels = c("food insecurity", "food security")
    ),
    race= factor(case_when(
      race == "1" ~ "Mexican American",  
      race == "2" ~ "Other Hispanic",
      race == "3" ~ "Non-Hispanic White",
      race == "4" ~ "Non-Hispanic Black",
      race == "6" ~ "Non-Hispanic Asian",
      race == "7" ~ "Other Race - Including Multi-Racial"),
      levels = c("Mexican American", "Other Hispanic","Non-Hispanic White","Non-Hispanic Black","Non-Hispanic Asian","Other Race - Including Multi-Racial")
    ),
    not_born_in_us=country_of_birth,
    education= factor(case_when(
      education == "1" ~ "Less than 9th grade",  
      education == "2" ~ "9-11th grade",
      education == "3" ~ "High school graduate/GED or equivalent",
      education == "4" ~ "Some college or AA degree",
      education == "5" ~ "College graduate or above"),
      levels = c("Less than 9th grade", "9-11th grade","High school graduate/GED or equivalent","Some college or AA degree","College graduate or above")
    ),
    marital_status=factor(case_when(
      marital_status == "1" ~ "Married/Living with Partner",  
      marital_status == "2" ~ "Widowed/Divorced/Separated",
      marital_status == "3" ~ "Never married"),
      levels = c("Married/Living with Partner", "Widowed/Divorced/Separated","Never married")
    ),
    family_income=poverty
  )
    
```
```{r, echo=FALSE}
dat %>% 
  select(food_security,gender,age,
    race,not_born_in_us,bmi,
    education,marital_status, family_income,
    hbp, diabetes,ckd,insurance) %>%
  tbl_summary(by=food_security) %>%
  bold_labels() %>%
  # modify_caption("**Table 1: Summary of Dataset**") %>%
  as_gt(include = everything()) %>%
  gt::tab_header("Table 1: Summary of Dataset by Food Security")
```
\end{singlespace}

The provided table summarizes characteristics of individuals divided into two groups based on food security status, with 3,163 facing food insecurity and 5,380 enjoying food security. A closer analysis reveals a similar gender distribution across both groups, a slightly younger median age in the food insecurity group (48 years) compared to the food security group (54 years), and marked differences in racial composition, notably with a higher percentage of Non-Hispanic Black individuals in the food insecurity group. A substantial proportion of the food insecure were not born in the U.S. (32%), which is higher than in the food secure group. Educational attainment varies significantly, with individuals having less education predominantly in the food insecurity group, while higher education levels are more common among those with food security. Marital status also differs, with a greater proportion of married individuals or those living with a partner in the food security group. Economic disparity is evident in the family income levels, with the food security group reporting higher median values. Health-wise, instances of high blood pressure and diabetes are slightly more prevalent in the food security group, although chronic kidney disease remains low in both. Finally, insurance coverage is more widespread in the food security group (89%) compared to the food insecurity group (76%), highlighting a possible link between economic stability and access to healthcare.

\newpage

## 3.2 Regression Analysis Results
```{r echo=FALSE, messgae=FALSE, warning=FALSE}
res1 <- read_csv("result_files/res_m1.csv")
```

```{r, echo=FALSE,include=FALSE}
res1 %>% 
  select(coef.names,OR,ci.low,ci.up) %>%
  knitr::kable()
```

\begingroup
\setstretch{1.0} 
\begin{center}
Table 2: Summary of Regression Coefficients with Odds Ratios and Confidence Intervals
\end{center}
|**coef names**  | **OR**|  **CI low**|  **CI up**|
|:----------------|---------:|---------:|---------:|
|Intercept        | 0.0865675| 0.0455321| 0.1645858|
|gender           |  |  |  |
|\quad male       | ref | ref | ref |
|\quad female     | 0.8882700| 0.7938057| 0.9939758|
|age              | 1.0242605| 1.0165347| 1.0320450|
|race      | | | |
|\quad Mexican American      | ref| ref| ref|
|\quad Other Hispanic      | 0.6944784| 0.4787410| 1.0074344|
|\quad Non-Hispanic White  | 1.6626392| 1.0989956| 2.5153595|
|\quad Non-Hispanic Black    | 1.1297815| 0.7313368| 1.7453055|
|\quad Non-Hispanic Asian    | 1.7310878| 1.1105557| 2.6983472|
|\quad Other Race - Including Multi-Racial    | 0.7473892| 0.3991244| 1.3995402|
|not born in U.S. | 1.1155619| 0.8022678| 1.5512006|
|bmi     | 0.9850796| 0.9752864| 0.9949711|
|education       | | | |
|\quad Less than 9th grade   | ref| ref| ref|
|\quad 9-11th grade| 1.3439491| 0.9764702| 1.8497227|
|\quad High school graduate/GED or equivalent  | 1.6825892| 1.2053256| 2.3488314|
|\quad Some college or AA degree | 2.1286519| 1.6312931| 2.7776485|
|\quad College graduate or above  | 3.7148680| 2.7390089| 5.0384081|
|marital status  | | | |
|\quad Married/Living with Partner  | ref| ref| ref|
|\quad Widowed/Divorced/Separated  | 0.7185600| 0.5343786| 0.9662221|
|\quad Never married | 0.8816866| 0.6728986| 1.1552577|
|family income  | 2.0249656| 1.8245303| 2.2474199|
|high blood pressure,    | 0.8545878| 0.6848649| 1.0663715|
|diabetes     | 0.7961340| 0.5730458| 1.1060710|
|chronic kidney disease      | 0.8875632| 0.5684104| 1.3859149|
|insurance   | 1.2391050| 0.9339267| 1.6440060|


\setstretch{2.0} 

The logistic regression analysis presented in the table identifies several significant predictors of food security, coded as 1 for food secure and 0 for food insecure individuals. Females are less likely to be food secure compared to males, with an odds ratio (OR) of 0.8882700, while age shows a positive association, with each additional year increasing the odds of food security (OR = 1.0242605). Racial disparities are evident, with Non-Hispanic Whites more likely to be food secure (OR = 1.6626392) compared to the reference group of Mexican Americans. Education level is a strong predictor, with higher education correlating to increased food security, particularly for college graduates or above (OR = 3.7148680). Marital status impacts food security, with married individuals or those living with a partner being the reference group; widowed, divorced, or separated individuals have lower odds of food security (OR = 0.7185600). Family income also plays a significant role; higher family income are associated with increased odds of food security (OR = 2.0249656). Health conditions like high blood pressure, diabetes, and chronic kidney disease are associated with lower odds of food security. Notably, having insurance is linked to higher odds of being food secure (OR = 1.2391050). These findings highlight the multifaceted nature of food security, influenced by a complex interplay of demographic, socioeconomic, and health factors.

## 3.3 Machine Learning Results


# 4. Conclusion



\newpage
# Reference
Rubin, D.B. (2018). Flexible Imputation of Missing Data, Second Edition. Chapman and Hall/CRC.

